---
title: "Pilot_A"
author: "Yochai Shavit"
date: "November 20, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = F)
```

The goal of this code is to make sure the "practice" code work from data as downloaded from Qualtrics, and make any code adjustments if necessary. 

```{r set wd, include=FALSE}
getwd()
```


#load packages
```{r packages}
library(knitr)
library(tidyverse)
library(ggplot2)
library(car)
library(ez)
library(broom)
library(pander)
```

#load data
```{r data}
df_old=read.csv("Pilot_A_older_raw.csv")
df_young=read.csv("Pilot_A_young_raw.csv")
```

```{r clean Qualtrics}
## df_old
colnames(df_old) <- as.character(unlist(df_old[1,]))
df_old=df_old[-c(1,2),]

#varaibles to remove
vars_remove=c('Start Date', 'End Date', 'Response Type', 'IP Address', 'Progress', 'Duration (in seconds)', 'Finished', 'Recorded Date', 'Recipient Last Name', 'Recipient First Name', 'Recipient Email','External Data Reference', 'Location Latitude', 'Location Longitude', 'Distribution Channel', 'User Language', 'issues_not_eligible', 'issues_not_interested', 'gender - Other: - Text', 'race - Other: - Text', 'issues_participant')

df_old=df_old%>%select(-one_of(vars_remove))

##df_young
colnames(df_young) <- as.character(unlist(df_young[1,]))
df_young=df_young[-c(1,2),]

df_young=df_young%>%select(-one_of(vars_remove))
```

```{r merge data sets}
df_og=full_join(df_old,df_young)

#Retain only participants who responded "yes" to bonus_interested, and passed attention check
df_og=df_og%>%filter(Bonus_eligible_interested=="Yes", attn_chk_20==20)

#retain only vars used in analysis
vars_remove=c('Bonus_eligible_interested', 'attn_chk_20')
df_og=df_og%>%select(-one_of(vars_remove))
```

#Tidy data- clean vars
```{r create rename vars, results='hide'}
#Add age group variable
df_og$age_grp=recode(df_og$age, "lo:35='young'; 60:hi='old'", as.factor.result = T)
summary(df_og$age_grp) #-> 3 of each

#Rename variables to be more readable
df_og=df_og%>%rename(subid=`Response ID`, gender=`gender - Selected Choice`, race=`race - Selected Choice`)

#Recode education to numeric
df_og$edu_lv=recode(df_og$education_level, "'Primary school or lower'=1; 'Middle school'=2; 'High school'=3; 'Undergraduate (College or equivalent)'=4; 'Graduate school or higher'=5", as.numeric.result = T)

#Recode subid into numbers (but still factor)
df_og$subid=as.factor(df_og$subid)
df_og$subid=as.numeric(df_og$subid)
df_og$subid=as.factor(df_og$subid)
str(df_og$subid)

#Make sure all variables that should be numeric are numeric
is.numeric(df_og$age); is.numeric(df_og$income); is.numeric(df_og$imp_relative_num); is.numeric(df_og$imp_friends_num); is.numeric(df_og$rel_1_mon); is.numeric(df_og$rel_1_time); is.numeric(df_og$nonrel_1_mon); is.numeric(df_og$nonrel_1_time)#-> All FALSE

#Recode all variables that should be numeric into numeric
df_og[,c(3,5:7,9:36)]=as.numeric(unlist(df_og[,c(3,5:7,9:36)])) #-> columns 3, 5:7 & 9:36 (all numeric vars)

#check
is.numeric(df_og$age); is.numeric(df_og$income); is.numeric(df_og$imp_relative_num); is.numeric(df_og$imp_friends_num); is.numeric(df_og$rel_1_mon); is.numeric(df_og$rel_1_time); is.numeric(df_og$nonrel_1_mon); is.numeric(df_og$nonrel_1_time)#-> all true!

```

#Descriptive demographics and statistics
Measure of central tendency and age group differences as reported in the original paper
```{r descriptives}
#table of descpritive statistics
dscrpt=df_og%>%group_by(age_grp)%>%summarise(male_female_ratio=paste(sum(gender=="Male"), "/", sum(gender=="Female")),
Age_av=round(mean(age),2),Age_sd=round(sd(age),2), 
Education_av=round(mean(edu_lv),2),Education_sd=round(sd(edu_lv),2), Household_income_av=round(mean(income),2),Household_income_sd=round(sd(income),2), sig_relatives_av=round(mean(imp_relative_num),2),sig_relatives_sd=round(sd(imp_relative_num),2), 
sig_friends_av=round(mean(imp_friends_num),2), sig_friends_sd=round(sd(imp_friends_num),2))

dscrpt2=dscrpt%>%rename('Age group'=age_grp, 'Male/Female ratio'=male_female_ratio)%>%unite('Age (m | sd)', c('Age_av', 'Age_sd'), sep=" | ")%>%unite('Education level (m | sd)', c('Education_av', 'Education_sd'), sep=" | ")%>%unite('Household income (m | sd)', c('Household_income_av', 'Household_income_sd'), sep=" | ")%>%unite('Important relatives (m | sd)', c('sig_relatives_av', 'sig_relatives_sd'), sep=" | ")%>%unite('Important friends (m | sd)', c('sig_friends_av', 'sig_friends_sd'), sep=" | ")

kable(dscrpt2,align='c', caption = "decsriptive statistics by age group") 

```

```{r test for differences}
#Gender (chi-square)
gen=chisq.test(df_og$gender, df_og$age_grp, rescale.p = T, correct = F)
#Age
age=t.test(data=df_og, age~age_grp)
#education level
edu_lv=t.test(data=df_og, edu_lv~age_grp)
#income level
inc_lv=t.test(data=df_og,income~age_grp)
#number of close relatives
rel=t.test(data=df_og, imp_relative_num~age_grp)
#number of close friends
frnd=t.test(data=df_og,imp_friends_num~age_grp)

pander(gen, digits=2, caption = "chi-square test for gender differences")
pander(age, digits=2, caption = "age group difference in average age")
pander(edu_lv, digits=2, caption = "age group difference in levels of education")
pander(inc_lv, digits=2, caption = "age group difference in average income")
pander(rel, digits=2, caption = "age group difference in average number of relatives")
pander(frnd, digits=2, caption = "age group difference in average number of friends")

```

#Tidy data- long form
data needs to be long-form for the first step, whereby each row is one trial for one subject (each subject has 14 rows (donation form * social distance), for each of the 2 kinship conditions (relative/nonrelative), resulting in 28 rows in total for 6 participants, meaning 168 observations. In addition, add age_group variable.
```{r tidy data df_og}
#long-form dataset
og_lng=df_og%>%
  gather(condition, amount,
         rel_1_mon,rel_2_mon,rel_5_mon,rel_10_mon,rel_20_mon, rel_50_mon, rel_100_mon,
         rel_1_time,rel_2_time,rel_5_time,rel_10_time,rel_20_time, rel_50_time, rel_100_time,
         nonrel_1_mon,nonrel_2_mon,nonrel_5_mon,nonrel_10_mon,nonrel_20_mon, nonrel_50_mon, nonrel_100_mon,
         nonrel_1_time,nonrel_2_time,nonrel_5_time,nonrel_10_time,nonrel_20_time, nonrel_50_time, nonrel_100_time
         )%>%
  separate(condition, c("kin", "soc_dist", "donation"), "_")%>%
  spread(kin, amount)%>%
  gather(kinship, amount,
         rel,nonrel) #-> That works!
```
