---
title: "practice_data"
author: "Yochai Shavit"
date: "October 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#set wd
```{r set wd}
getwd()
setwd("~/Stanford Ph.D/3rd year/Psych 251/Replication project/shavit2017/practice_analysis")
```

#load packages
```{r load packages}
library(knitr)
library(tidyverse)
library(lavaan)
library(car)
library(broom)
library(AUC)
```

#create data 
First, create dataset as it will be downloaded from Qualtrix (wide form with raw values of time and money in each of the conditions), including demographic data. This is to test if I can fit the hyperbolic discount function in R

```{r create og practice df, echo=FALSE}
#Create dataframe with values attempting to be close to the ones in the paper
df=data.frame(row.names = c(1:108))
  df$subid=c(1:108)
  df$age=ifelse(df$subid<=54, sample(x=c(60:89), size = 54, replace = T), sample(x=c(18:35), size = 54, replace = T))
  
  df$rel_1_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (102.08/(1+0.32*1)), sd=20), rnorm(n=54, mean=(92.96/(1+0.19*1)), sd=20))
  df$rel_1_time=ifelse(df$subid<=54, rnorm(n=54, mean = (91.35/(1+0.33*1)), sd=20), rnorm(n=54, mean=(48.44/(1+0.43*1)), sd=20))
  
  df$rel_2_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (102.08/(1+0.32*2)), sd=20), rnorm(n=54, mean=(92.96/(1+0.19*2)), sd=20))
  df$rel_2_time=ifelse(df$subid<=54, rnorm(n=54, mean = (91.35/(1+0.33*2)), sd=20), rnorm(n=54, mean=(48.44/(1+0.43*2)), sd=20))
  
  df$rel_5_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (102.08/(1+0.32*5)), sd=20), rnorm(n=54, mean=(92.96/(1+0.19*5)), sd=20))
  df$rel_5_time=ifelse(df$subid<=54, rnorm(n=54, mean = (91.35/(1+0.33*5)), sd=20), rnorm(n=54, mean=(48.44/(1+0.43*5)), sd=20))
  
  df$rel_10_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (102.08/(1+0.32*10)), sd=20), rnorm(n=54, mean=(92.96/(1+0.19*10)), sd=20))
  df$rel_10_time=ifelse(df$subid<=54, rnorm(n=54, mean = (91.35/(1+0.33*10)), sd=20), rnorm(n=54, mean=(48.44/(1+0.43*10)), sd=20))
  
  df$rel_20_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (102.08/(1+0.32*20)), sd=20), rnorm(n=54, mean=(92.96/(1+0.19*20)), sd=20))
  df$rel_20_time=ifelse(df$subid<=54, rnorm(n=54, mean = (91.35/(1+0.33*20)), sd=20), rnorm(n=54, mean=(48.44/(1+0.43*20)), sd=20))
  
  df$rel_50_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (102.08/(1+0.32*50)), sd=20), rnorm(n=54, mean=(92.96/(1+0.19*50)), sd=20))
  df$rel_50_time=ifelse(df$subid<=54, rnorm(n=54, mean = (91.35/(1+0.33*50)), sd=20), rnorm(n=54, mean=(48.44/(1+0.43*50)), sd=20))
  
  df$rel_100_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (102.08/(1+0.32*100)), sd=20), rnorm(n=54, mean=(92.96/(1+0.19*100)), sd=20))
  df$rel_100_time=ifelse(df$subid<=54, rnorm(n=54, mean = (91.35/(1+0.33*100)), sd=20), rnorm(n=54, mean=(48.44/(1+0.43*100)), sd=20))
  
  df$nonrel_1_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (12.48/(1+0.23*1)), sd=20), rnorm(n=54, mean=(93.35/(1+0.16*1)), sd=20))
  df$nonrel_1_time=ifelse(df$subid<=54, rnorm(n=54, mean = (14.22/(1+0.28*1)), sd=20), rnorm(n=54, mean=(47.29/(1+0.5*1)), sd=20))
  
  df$nonrel_2_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (12.48/(1+0.23*2)), sd=20), rnorm(n=54, mean=(93.35/(1+0.16*2)), sd=20))
  df$nonrel_2_time=ifelse(df$subid<=54, rnorm(n=54, mean = (14.22/(1+0.28*2)), sd=20), rnorm(n=54, mean=(47.29/(1+0.5*2)), sd=20))
  
  df$nonrel_5_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (12.48/(1+0.23*5)), sd=20), rnorm(n=54, mean=(93.35/(1+0.16*5)), sd=20))
  df$nonrel_5_time=ifelse(df$subid<=54, rnorm(n=54, mean = (14.22/(1+0.28*5)), sd=20), rnorm(n=54, mean=(47.29/(1+0.5*5)), sd=20))
  
  df$nonrel_10_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (12.48/(1+0.23*10)), sd=20), rnorm(n=54, mean=(93.35/(1+0.16*10)), sd=20))
  df$nonrel_10_time=ifelse(df$subid<=54, rnorm(n=54, mean = (14.22/(1+0.28*10)), sd=20), rnorm(n=54, mean=(47.29/(1+0.5*10)), sd=20))
  
  df$nonrel_20_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (12.48/(1+0.23*20)), sd=20), rnorm(n=54, mean=(93.35/(1+0.16*20)), sd=20))
  df$nonrel_20_time=ifelse(df$subid<=54, rnorm(n=54, mean = (14.22/(1+0.28*20)), sd=20), rnorm(n=54, mean=(47.29/(1+0.5*20)), sd=20))
  
  df$nonrel_50_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (12.48/(1+0.23*50)), sd=20), rnorm(n=54, mean=(93.35/(1+0.16*50)), sd=20))
  df$nonrel_50_time=ifelse(df$subid<=54, rnorm(n=54, mean = (14.22/(1+0.28*50)), sd=20), rnorm(n=54, mean=(47.29/(1+0.5*50)), sd=20))
  
  df$nonrel_100_mon=ifelse(df$subid<=54, rnorm(n=54, mean = (12.48/(1+0.23*100)), sd=20), rnorm(n=54, mean=(93.35/(1+0.16*100)), sd=20))
  df$nonrel_100_time=ifelse(df$subid<=54, rnorm(n=54, mean = (14.22/(1+0.28*100)), sd=20), rnorm(n=54, mean=(47.29/(1+0.5*100)), sd=20))
  
  df$incm_lv=ifelse(df$subid<=54, rnorm(n=54, mean = 0.82, sd=0.59), rnorm(n=54, mean=0.94, sd=0.5))
  df$edu_lv=ifelse(df$subid<=54, sample(x=c(1:5), size=54, replace = T), sample(x=c(1:5), size = 54, replace = T))
  df$num_friends=ifelse(df$subid<=54, rnorm(n=54, mean = 10.08, sd=9.96), rnorm(n=54, mean=8.10, sd=6.97))
  df$num_relatives=ifelse(df$subid<=54, rnorm(n=54, mean = 12.36, sd=11.83), rnorm(n=54, mean=8.98, sd=6.41))

```
 
 This dataset is obviously full of impossible values (negative number of friends or value of donations). Would it make sense to set all negative values to zero? In the paper they mention fitting the curve to the group median data because it is typically not normally distributed, so using a normal distribution here probabky wouldn't be different because the mean would be the same as the median.

Let's check this:
```{r summary raw values, echo=FALSE}
summary(df$rel_1_mon);summary(df$rel_2_mon);summary(df$rel_5_mon);summary(df$rel_10_mon);summary(df$rel_20_mon);summary(df$rel_50_mon);summary(df$rel_100_mon) #-> means and medians are quite close

summary(df$rel_1_time);summary(df$rel_2_time);summary(df$rel_5_time);summary(df$rel_10_time);summary(df$rel_20_time);summary(df$rel_50_time);summary(df$rel_100_time) #-> a few have quite a difference (over 1.5 in value)

summary(df$nonrel_1_mon);summary(df$nonrel_2_mon);summary(df$nonrel_5_mon);summary(df$nonrel_10_mon);summary(df$nonrel_20_mon);summary(df$nonrel_50_mon);summary(df$nonrel_100_mon)#-> a few have quite a difference (over 1.5 in value)

summary(df$nonrel_1_time);summary(df$nonrel_2_time);summary(df$nonrel_5_time);summary(df$nonrel_10_time);summary(df$nonrel_20_time);summary(df$nonrel_50_time);summary(df$nonrel_100_time)#-> a few have quite a difference (over 1.5 in value)
```

#create second dataset, with random sampling from a certain range of values

```{r og dataset, echo=FALSE}
#Create new dataset with plausible raw values as will be downloaded from qualtrics (decrease 0.5*distance from range with each move of 1 social distance)
df_og=data.frame(row.names = c(1:108))
  df_og$subid=c(1:108)
  df_og$age=ifelse(df_og$subid<=54, sample(x=c(60:89), size = 54, replace = T), sample(x=c(18:35), size = 54, replace = T))
  
  df_og$rel_1_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(75:100), replace = T), sample(size=54, x=c(70:100), replace = T))
  df_og$rel_1_time=ifelse(df_og$subid<=54, sample(size=54, x = c(75:100), replace = T), sample(size=54, x=c(40:70), replace = T))
  
  df_og$rel_2_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(74:99), replace = T), sample(size=54, x=c(69:99), replace = T))
  df_og$rel_2_time=ifelse(df_og$subid<=54, sample(size=54, x = c(74:99), replace = T), sample(size=54, x=c(39:69), replace = T))
  
  df_og$rel_5_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(72.5:97.5), replace = T), sample(size=54, x=c(67.5:97.5), replace = T))
  df_og$rel_5_time=ifelse(df_og$subid<=54, sample(size=54, x = c(72.5:97.5), replace = T), sample(size=54, x=c(37.5:67.5), replace = T))
  
  df_og$rel_10_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(70:95), replace = T), sample(size=54, x=c(65:95), replace = T))
  df_og$rel_10_time=ifelse(df_og$subid<=54, sample(size=54, x = c(70:95), replace = T), sample(size=54, x=c(35:65), replace = T))
  
  df_og$rel_20_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(65:90), replace = T), sample(size=54, x=c(60:90), replace = T))
  df_og$rel_20_time=ifelse(df_og$subid<=54, sample(size=54, x = c(65:90), replace = T), sample(size=54, x=c(30:60), replace = T))
  
  df_og$rel_50_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(50:75), replace = T), sample(size=54, x=c(45:75), replace = T))
  df_og$rel_50_time=ifelse(df_og$subid<=54, sample(size=54, x = c(50:75), replace = T), sample(size=54, x=c(15:45), replace = T))
  
  df_og$rel_100_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(25:50), replace = T), sample(size=54, x=c(20:50), replace = T))
  df_og$rel_100_time=ifelse(df_og$subid<=54, sample(size=54, x = c(25:50), replace = T), sample(size=54, x=c(0:20), replace = T))
  
  df_og$nonrel_1_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(40:50), replace = T), sample(size=54, x=c(70:100), replace = T))
  df_og$nonrel_1_time=ifelse(df_og$subid<=54, sample(size=54, x = c(40:50), replace = T), sample(size=54, x=c(40:70), replace = T))
  
  df_og$nonrel_2_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(39:49), replace = T), sample(size=54, x=c(69:99), replace = T))
  df_og$nonrel_2_time=ifelse(df_og$subid<=54, sample(size=54, x = c(39:49), replace = T), sample(size=54, x=c(39:69), replace = T))
  
  df_og$nonrel_5_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(37.5:47.5), replace = T), sample(size=54, x=c(67.5:97.5), replace = T))
  df_og$nonrel_5_time=ifelse(df_og$subid<=54, sample(size=54, x = c(37.5:47.5), replace = T), sample(size=54, x=c(37.5:67.5), replace = T))
  
  df_og$nonrel_10_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(35:45), replace = T), sample(size=54, x=c(65:95), replace = T))
  df_og$nonrel_10_time=ifelse(df_og$subid<=54, sample(size=54, x = c(35:45), replace = T), sample(size=54, x=c(35:65), replace = T))
  
  df_og$nonrel_20_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(30:40), replace = T), sample(size=54, x=c(60:90), replace = T))
  df_og$nonrel_20_time=ifelse(df_og$subid<=54, sample(size=54, x = c(30:40), replace = T), sample(size=54, x=c(30:80), replace = T))
  
  df_og$nonrel_50_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(15:25), replace = T), sample(size=54, x=c(45:75), replace = T))
  df_og$nonrel_50_time=ifelse(df_og$subid<=54, sample(size=54, x = c(15:25), replace = T), sample(size=54, x=c(15:45), replace = T))
  
  df_og$nonrel_100_mon=ifelse(df_og$subid<=54, sample(size=54, x = c(0:10), replace = T), sample(size=54, x=c(20:50), replace = T))
  df_og$nonrel_100_time=ifelse(df_og$subid<=54, sample(size=54, x = c(0:10), replace = T), sample(size=54, x=c(0:20), replace = T))
  
  df_og$incm_lv=ifelse(df_og$subid<=54, rnorm(n=54, mean = 0.82, sd=0.59), rnorm(n=54, mean=0.94, sd=0.5))
  df_og$edu_lv=ifelse(df_og$subid<=54, sample(x=c(1:4), size=54, replace = T), sample(x=c(3:5), size = 54, replace = T))
  df_og$num_friends=ifelse(df_og$subid<=54, rnorm(n=54, mean = 10.08, sd=9.96), rnorm(n=54, mean=8.10, sd=6.97))
  df_og$num_relatives=ifelse(df_og$subid<=54, rnorm(n=54, mean = 12.36, sd=11.83), rnorm(n=54, mean=8.98, sd=6.41))
  df_og$gender=sample(x=c("female", "male"), size = 108, replace=T)
  #Add age group variable
df_og$age_grp=recode(df_og$age, "lo:35='young'; 60:hi='old'", as.factor.result = T)
summary(df_og$age_grp) #-> 54 of each, like I generated the data

```

#Descriptive demographics and statistics
Measure of central tendency and age group differences as reported in the original paper
```{r descriptives}
#table of descpritive statistics
dscrpt=df_og%>%group_by(age_grp)%>%summarise(male_female_ratio=paste(sum(gender=="male"), "/", sum(gender=="female")),
Age_av=mean(age),Age_sd=sd(age), 
Education_av=mean(edu_lv),Education_sd=sd(edu_lv), Household_income_av=mean(incm_lv),Household_income_sd=sd(incm_lv), sig_relatives_av=mean(num_relatives),sig_relatives_sd=sd(num_relatives), 
sig_friends_av=mean(num_friends), sig_friends_sd=sd(num_friends))

kable(dscrpt, digits = 2, align='c', caption = "decsriptive statistics by age group")
```

```{r test for differences}
#Gender (chi-square)
with(df_og,
print(chisq.test(gender, age_grp, rescale.p = T, correct = F)))
#Age
with(df_og,
print(t.test(age~age_grp)))
#education level
with(df_og,
print(t.test(edu_lv~age_grp)))
#income level
with(df_og,
print(t.test(incm_lv~age_grp)))
#number of close relatives
with(df_og,
print(t.test(num_relatives~age_grp)))
#number of close friends
with(df_og,
print(t.test(num_friends~age_grp)))

```

**Note-** These data serve, to some extent, as a simulation of the dempgraphic effects in the original paper. The fact that the simulated values for number of close friends and relatives is sometime significant and sometimes not (for both) suggests that the finding that age groups differed in number of relatives but not in number of friends is at the very least fragile. 

#Tidy data
data needs to be long-form for the first step, whereby each row is one trail for one subject (each subject has 14 rows (donation form * social distance), for each of the 2 kinship conditions (relative/nonrelative), resulting in 28 rows in total for 108 participants, meaning 3024 observations. In addition, add age_group variable.
```{r tidy data df_og}

#long-form dataset
og_lng=df_og%>%
  gather(condition, amount,
         rel_1_mon,rel_2_mon,rel_5_mon,rel_10_mon,rel_20_mon, rel_50_mon, rel_100_mon,
         rel_1_time,rel_2_time,rel_5_time,rel_10_time,rel_20_time, rel_50_time, rel_100_time,
         nonrel_1_mon,nonrel_2_mon,nonrel_5_mon,nonrel_10_mon,nonrel_20_mon, nonrel_50_mon, nonrel_100_mon,
         nonrel_1_time,nonrel_2_time,nonrel_5_time,nonrel_10_time,nonrel_20_time, nonrel_50_time, nonrel_100_time
         )%>%
  separate(condition, c("kin", "soc_dist", "donation"), "_")%>%
  spread(kin, amount)%>%
  gather(kinship, amount,
         rel,nonrel) #-> That works!
```

#Curve fitting
Now that the dataset is tidy, proceed to attemp and fit the hyperbolic discount curve to the median value in each group (age_group-kinship-donation form)
```{r median values}
med_table=og_lng%>%group_by(age_grp,kinship,donation,soc_dist)%>%summarise(med_amnt=median(amount))
write_csv(med_table,"med_table_shavit.csv")

ggplot(data=med_table, aes(x=soc_dist, y=med_amnt, color=age_grp))+
  geom_point()+
  facet_grid(donation~kinship)

#for each age_grp*kinship*donation condition
old_rel_mon=med_table%>%filter(age_grp=="old" & kinship=="rel" & donation=="mon")
kable(old_rel_mon, digits = 2, align = 'c', caption = "Older,relative,money: median amount given")

old_rel_time=med_table%>%filter(age_grp=="old" & kinship=="rel" & donation=="time")
kable(old_rel_time, digits = 2, align = 'c', caption = "Older,relative,time: median amount given")

yng_rel_mon=med_table%>%filter(age_grp=="young" & kinship=="rel" & donation=="mon")
kable(yng_rel_mon, digits = 2, align = 'c', caption = "Younger,relative,money: median amount given")

yng_rel_time=med_table%>%filter(age_grp=="young" & kinship=="rel" & donation=="time")
kable(yng_rel_time, digits = 2, align = 'c', caption = "Younger,relative,time: median amount given")

old_nonrel_mon=med_table%>%filter(age_grp=="old" & kinship=="nonrel" & donation=="mon")
kable(old_nonrel_mon, digits = 2, align = 'c', caption = "Older,non-relative,money: median amount given")

old_nonrel_time=med_table%>%filter(age_grp=="old" & kinship=="nonrel" & donation=="time")
kable(old_nonrel_time, digits = 2, align = 'c', caption = "Older,non-relative,time: median amount given")

yng_nonrel_mon=med_table%>%filter(age_grp=="young" & kinship=="nonrel" & donation=="mon")
kable(yng_nonrel_mon, digits = 2, align = 'c', caption = "Younger,non-relative,money: median amount given")

yng_nonrel_time=med_table%>%filter(age_grp=="young" & kinship=="nonrel" & donation=="time")
kable(yng_nonrel_time, digits = 2, align = 'c', caption = "Younger,non-relative,time: median amount given")
```

#Code from Mike
```{r}
#define dataset
d=med_table
d$soc_dist=as.numeric(d$soc_dist)

# first plot
ggplot(d, 
       aes(x = soc_dist, y = med_amnt, col = age_grp)) + 
  geom_point() + 
  facet_grid(kinship ~ donation) + 
  geom_smooth(method="lm", formula = y ~ poly(x, 2), 
              se = FALSE)

# for testing
data_subset <- filter(d, age_grp == "old", 
                      kinship == "nonrel", 
                      donation == "mon")

# v=V/(1 +kN)
# where V is the undiscounted reward value, 
# v is the reward value that a person would like to forgo for the benefit for someone at a social distance N, and k is a constant indexing degree of social discounting (i.e., discount rate) across social distances.

# where R-square and root-mean-square error (RMSE) were used to assess the goodness of  t. 

# compute the hyperbolic function
discount <- function (V, k, N) {
  nu <- V / (1 + k * N)
  
  return(nu)
}

# compute RMSE for a given parameter set and data
data_vs_discount <- function(V, k, N, amount) {
  nu <- discount(V, k, N)
  
  RMSE <- sum(sqrt((amount - nu)^2))
  
  return(RMSE)
}

# wrapper to feed to optim
optim_discount_wrapper <- function(x) {
  return(data_vs_discount(V = x[1], k = x[2],  
                          N = data_subset$soc_dist, 
                          amount = data_subset$med_amnt))
}

# try fitting optimizer
opt1=optim(c(100, .2), optim_discount_wrapper)
opt1 #-> first par= V, second par=k

# get predictions on test case
data_subset$pred <- discount(V = 46, k = .02,
                             N = data_subset$soc_dist)    

## run for each age group, donation and kinship combination ##
# check out the `purrr` package
d %>%
  group_by(age_grp, kinship, donation) %>%
  do(optim(c(100, .2), function(x) {
    return(data_vs_discount(V = x[1], k = x[2],  
                            N = .$soc_dist, 
                            amount = .$med_amnt))
  }))
```

# finding AUC
```{r}
# first step - using Erin calculous calculator (and WolframAlpha), the integral of this discount function is:
# V/k*ln(abs(1+k*n))+c -> we don't really care about c for AUC.

# define V and k from the code above
k=opt1$par[2]
V=opt1$par[1]

#Evaluate for upper bound
up=V/k*log(abs(1+k*100))
lower=V/k*log(abs(1+k*0))

AUC=up-lower
log(AUC)
```

