---
title: "Replication of Study 1 by Gong, Zhang & Fung (2017, Journals of Gerontology: Series B)"
author: "Yochai Z. Shavit (yochais@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---

##Introduction

  Socioemotional Selectivity Theory (SST- Carstensen, 2006) postulates that due to a sense of limited time left in life older adults are motivated to engage with meaningful activities over ones that might provide future benefits. Yet, despite being thought of as a meaningful activity (Greenfield & Marks, 2004), older adults do not volunteer in higher rates compared to younger cohorts (U.S Bureau of Labor, 2015). The relatively low volunteer rates among older adults is especially surprising since laboratory experiments found that older adults show more prosocial behaviors compared to younger adults, and are more concerned with the wellbeing of others in need (Sze, Gyurak, Goodkind, &  Levenson, 2012; Beadle, Sheehan, Dahlben, & Gutchess, 2013;  Hubbard, Harbaugh, Srivastava, Degras, & Mayr, 2016)


  The study I attempt to replicate by Gong, Zhang, & Fung (2017) provides a possible explanation for this apparent discrepancy. In a hypothetical donation task, Gong and colleagues asked younger (N=89) and older (N=66) Hong Kong Chinese community membersto indicate how much money and time they are willing to donate to a person in need, with this person being progressively more socially distant to them. They found: a *weak qualified* main effect ($\eta_p^2$=0.05) of age such that older adults were less generous than younger adults when averaging across donation (money or time) and kinship (relative or non-relative) types. However, they found a 3-way Age group X Donation X Kinship interaction showing that *when donating their time to relatives, older adults were more generous than younger adults*. This post-hoc test had a medium effect size (*Cohen's D*=`r round(((2*3.22)/sqrt(145)),2)`, calculated based on reported t-statistic of $t_{(145)}$=3.22). **the qualified main effect of age group and the post-hoc test suggesting that older adults are more generous than younger adults when donating time to relatives are the two target findings for this replication**. 
  
  In addition, their post-hoc comparisons revealed that older adults, compared to younger adults, were less generous to non-relatives in monetary donations but the groups did not differ in generousity towards relatives in this donation condition. In contrast, older adults were less generous than younger adults when donating time to non-relatives (but were more generous to relatives as noted above). 
  
  Gong, Zhang and Fung were also interested in comparing the social discount rate of older and younger adults in each of the conditions. They found no main effects of social discounting, and an Age group X Donation type interaction. Post-hoc tests revealed that older and younger adults' social discount rates did not differ when donating time, and that compared to younger adults older adults had higher discount rates when donating money to relatives and non-relatives alike. They conclude that their findings regarding donation amounts and social discount rates are in accord with SST- older adults are affected more than younger adults by kinship and degree of social closeness, though these effects depend on the context of giving.

##Methods

###Power Analysis

The original effect sizes of interest were: $\eta_p^2$=0.05 for the qualified main age group effect effect, and *D*=`r round(((2*3.22)/sqrt(145)),2)` for the post-hoc t-test on generosity of time donations for relatives. 

Power analysis using G*Power version 3.1.2.9 suggested that in order to obtain 80%, 90% and 95% power for the main effect of age, the required sample sizes are 154, 204, and 252 respectively (details can be found here: [Age group main effect- 80% power](age_grp_main_pwr_80.png), [Age group main effect- 90% power](age_grp_main_pwr_90.png), and [Age group main effect- 95% power](age_grp_main_pwr_95.png)). 

In order to obtain 80%, 90%, and 95% power to observe the stronger effect of age group on generosity in time donations to relatives (Older>Younger), requires sample sizes of 90 (45 per group), 124 (62 per group), and 156 (78 per group) respectively (details can be found here: [age group time donations, relatives- 80% power](t_test_pwr_80.png), [age group time donations, relatives- 90% power](t_test_pwr_90.png), and [age group time donations, relatives- 95% power](t_test_pwr_95.png)).

**Feasability-** Since the task requires about 10 minutes to complete online, and since the required samples are somewhat large, the replication study will be powered to observe the stronger effect of age group on generosity in time donations to relatives at 80% power.

###Planned Sample

The above mentioned power calculation suggests that a sample consisting of **45 older adults** (aged 60 or older), and **45 younger adults** (aged 35 or younger) will suffice to achieve 80% to detect the effect of interest. We will rely on Amazon's mechanical Turk for data colelction, and sample only U.S residents to minimize the chance that participants do not read English.

###Materials

In this study, age group is treated as a between subjects independent variable. Participants in both age groups completed the same experimental task. At first, the were given the following instruction:

> "The following task asks you to imagine that you have a list of 100 relatives (or non-relatives) arranged in descending order of their closeness with you. The person at position #1 is your closest relative (or non-relative), while the person at position #100 is a relative (or nonrelative) you may know but are not close at all. You do not have to actually create the list — just imagine that you have done so. Next you will be asked to answer questions regarding these relatives (or non-relatives) at a given position."

Participants were then asked to indicate how much money and time they would be willing to hypothetically donate to a person in a given position (N) on that list (N=1,2,5,10,20,50, and 100). The promt read as follows:

> "Imagine that the relative (or non-relative) at position N on your list is now hospitalized. He/she needs to be accompanied and taken care of in the hospital, as well as a large amount of money to pay for the medical expense. If you now own 100 thousand of Hong Kong dollars and 100 days of vacation, how much money would you donate to him/her, and how many days would you spend taking care of him/her?"

These materials are used verbatim in the replication attempt, with the exception of replacing "Hong Kong dollars" with "U.S dollars". The original authors' advice was sought in order to decide whether to change the amount to the $U.S equivalent of 100 thousand Hong Kong dollars (~18 thousand U.S dollars). The first author indicated that it is better to retain the value of 100 thousand and change only the currency.

###Procedure	

Participants first provided demographic information regarding their age, gender, level of education and household income. In addition, participants were asked to "*directly estimate the number of important relatives/ friends*" that they have, since this was seen as an important control variable.

Participants then completed the Chinese version of the Alzheimer’s Disease Assessment Scale-Cognitive Subscale (ADAS-Cog). Since this is an in-person assesment, **this stage is not followed in the replication study**. 

Participants then proceeded to the hypothetical donation task reported above. 

**For the replication attempt**- In order to recruit participants on Amazon's Mechanical Turk, two surveys were set on using the Qualtrics platform. one survey was desigend to allow only people who indicate their age as being over 60 to take the survey, while the other was designed to allow only those who indicate their age as being under 35 to proceed to the survey. Participants who accept the HIT, are compensated \$0.1 for taking the short demographic survey (including age and the demographic variables detailed above), and could opt to recieve a bonus payment of \$1.21 by completeing the experimental task. 

**links to working surveys**: 

[Older adults](https://stanforduniversity.qualtrics.com/jfe/form/SV_1G5Gp4eys77yAHr)

[Younger adults](https://stanforduniversity.qualtrics.com/jfe/form/SV_1QYqUm51SjqaQfP) 

###Analysis Plan
As a result of the replication study being done online, the two datasets (one with data from older participants and another with younger participants' data) first have to be merged. After this merger, the analysis proceeds as follows:

The analysis is based on fitting hyperbolic discount curves for each individual in each of the 2(donation form) X 2(kinship) conditions. This is done first on the group median values to assess the overall appropriateness of the function. 

> "To analyze the donation data, we first modeled the group donation data. For each age × donation form × kinship cell, a hyperbolic discount function (i.e., Equation 1) was fit to the group median values (Jones & Rachlin, 2006; Rachlin & Jones, 2007, 2008) over the seven social distances (N = 1, 2, 5, 10, 20, 50, and 100):

> $\nu$ = V / (1 + $\kappa$N) $_{(1)}$

> where V is the undiscounted reward value, v is the reward value that a person would like to forgo for the benefit for someone at a social distance N, and k is a constant indexing degree of social discounting (i.e., discount rate) across social distances. Following the convention of the literature, group median values instead of mean values were used"

After assuring that the hypoerbolic discount curve fit the group data well by computing the Root Mean Square Error (RMSE) and variance explaines ($R^2$), the same curve is fit for each participant, extracting the values for Area under the curve (*AUC*) as an indicator of overall generosity, and discount rate ($\kappa$):

> "Once the appropriateness of the hyperbolic discount function was confirmed for group data, we then conducted the same fitting process to individual data to get the discount rate ($\kappa$) and the *AUC* for each participant under different conditions. Again following the convention of the literature, individual data that did not fit the hyperbolic discount function were excluded from further analyses (Vuchinich & Simpson, 1998). The individuals’ $\kappa$ and *AUC* values were then normalized by a natural logarithm transformation (e.g., Margittai et al., 2015), and transformed values that exceeded ±3 SDs were treated as outliers."

The individual *AUC* and $\kappa$ values are then entered to a dataset comprising of 4 rows per participant corresponding to the four 2X2 cells. To analyze these data:

> "the transformed individual *AUC* values and $\kappa$ values were submitted respectively to a 2 (age: younger and older) × 2 (donation form: money and time) × 2 (kinship: relatives and nonrelatives) mixed design, repeated measures ANOVA (and post hoc tests), with and without controlling for demographic variables"

**Key analysis of interest-** The mixed design repeated measure ANOVA of *AUC* values is the main interest for the purpose of this replication, with the post hoc test comparing older and younger adults' *AUC* values when donating time to relatives. 

"Main" effect of age will be taken as the affect of Age when controlling for all other main independent variables and their interactions (type III sum of squares, as is typically used in Statistical programs). The authors did not specify the type of post-hoc test used to examine group differences in each condition. Since there are relatively few (4) pairwise comparisons in this test, and out of desire to be cautious, I will use the Bonferroni correction to deterimine the statistical significance of this test.
Though not of main interest for this replication, I will also run the same ANOVA and post-hoc procedures on individual $\kappa$ values.


###Differences from Original Study

This replication is different form the original study in a number of ways: 

First, the original study sampled Honk-Kong Chinese, which might have different prosocial behavior patterns than U.S citizens. However, the authors did not state in their study that their result should only extend to a particular cultural/ethnic group. 

Second, as noted above, the amount of money participants are instructed to imagine having is different in relative terms. 100 thousand Hong Kong dollars are considerably less valuable (in terms of purchase power) then 100 thousand U.S dollars, even when adjusting to the difference in cost of living between the two countries. However, as stated, the primary author of the original paper instructed the author of this replication to keep the value to 100 thousand and change only the currency. 

Third, the original study was conducted in lab and included a cognitive assesment procedure. The replication is taking place on Amazon's mechanical Turk and does not include cognitive assessment. The decision to drop this assesment was made since the original paper did not find a relationship between performance and cognitive ability, and the first author of the original paper stated in a personal communication that he does not think that this procedure would alter participants' mindsets. 


```{r include=FALSE}
### Methods Addendum (Post Data Collection)
#You can comment this section out prior to final report with data collection.
```


#### Actual Sample
  In Pilot A, reported here, responses of `r 6+14` potential participants were recorded. In order to pilot the data cleaning procedures, some of the participants purposefully provided answers that rendered them ineligible. The final sample consisted of 3 non-naive older participants and 3 non-naive younger participants.

#### Differences from pre-data collection methods plan
Since this pilot was not done through Amazon Mechanical Turk, possible issues with interacting with that platform for not tested.


##Results


### Data preparation

Data preparation following the analysis plan.
	
```{r include=F}
###Data Preparation

####Load Relevant Libraries and Functions

####Import data

#### Data exclusion / filtering

#### Prepare data for analysis - create columns etc.
```

### Confirmatory analysis

The analyses as specified in the analysis plan.  

*Side-by-side graph with original graph is ideal here*

###Exploratory analyses

Any follow-up analyses desired (not required).  

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.
